### 2012-02-13 docs/SDC/sdc.tex
new version of reaction term for predicting h

\subsubsection{Species Source Terms.}
For the species prediction, the form of $I_{\rho X_k}$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}):  
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

(Andy's Idea) Define $I_{X_k}$ using
\begin{equation}
I_{X_k} = \frac{X_k^\mathrm{new} - X_k^\mathrm{old}}{\Delta t} - A_{X_k},
\end{equation}
where we first define a state that has only been updated with advection:
\begin{equation}
\frac{(\rho X_k)^{(1)} - (\rho X_k)^\mathrm{old}}{\Delta t} = A_{\rho X_k},
\end{equation}
and then define the species mass fractions,
\begin{equation}
X_k^{(1)} = (\rho X_k)^{(1)} / \sum_k (\rho X_k)^{(1)}, \quad 
X_k^\mathrm{old} = (\rho X_k)^\mathrm{old} / \sum_k (\rho X_k)^\mathrm{old}, \quad 
X_k^\mathrm{new} = (\rho X_k)^\mathrm{new} / \sum_k (\rho X_k)^\mathrm{new},
\end{equation}
and finally define $A_{X_k}$ using
\begin{equation}
\frac{X^{(1)} - X^\mathrm{old}}{\Delta t}= A_{X_k}.
\end{equation}

\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
%
\begin{equation}
I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k} 
\label{eq:sdc:Irhoo}
\end{equation}

\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)
%
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ]
\end{equation}

We note that there is no $I$ term for $\rho$ or $\rho'$ prediction, since
the density evolution equation does not have a reaction source term.

@@

\begin{itemize}
\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhoprime}

Not implemented.

Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
equation (see Eq.~\ref{rhohprime equation}).  Defining the source
terms for this requires a base state enthalpy that knows about the
source terms---this is currently not implemented.

In particular, should $(\rho h)_0$ see a $\overline{(\rho \Hnuc)}$ term?
what about an average thermal diffusion?

@@

This is the most straightforward prediction type.  The SDC solver
integrates the equation for $(\rho h)$:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt}  + \rho H_{\rm nuc} 
\end{equation}
(shown here without diffusion or external heat sources).  Expanding
the time derivative and divergence, and using the continuity equation
we see:
\begin{equation}
\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho} \frac{Dp_0}{Dt}  + \frac{1}{\rho} (\rho H_{\rm nuc}) \label{eq:sdc:h}
\end{equation}
Comparing these equations, we see that 
\begin{equation}
I_{h}  = \frac{1}{\rho^{n+\myhalf}} \left [
    \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
\end{equation}

>>>>

\subsubsection{Species Source Terms.}
For the species prediction, the form of $I$ depends on
{\tt species\_pred\_type} (see \S \ref{sec:pred:density}).
We note that there is no $I$ term for $\rho$ or $\rho'$ prediction, since
the density evolution equation does not have a reaction source term.
\begin{itemize}
\item {\tt species\_pred\_type} = 1 (predict $\rho'$ and $X_k$)
\begin{equation}
I_{X_k} = \frac{1}{\rho^{n+\myhalf}} \left [ 
      \frac{(\rho X_k)^\mathrm{new} - 
            (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}  \right ].
\end{equation}
(Andy's Idea) Define $I_{X_k}$ using
\begin{equation}
I_{X_k} = \frac{X_k^\mathrm{new} - X_k^\mathrm{old}}{\Delta t} - A_{X_k},
\end{equation}
where we first define a state that has only been updated with advection:
\begin{equation}
\frac{(\rho X_k)^{(1)} - (\rho X_k)^\mathrm{old}}{\Delta t} = A_{\rho X_k},
\end{equation}
and then define the species mass fractions,
\begin{equation}
X_k^{(1)} = (\rho X_k)^{(1)} / \sum_k (\rho X_k)^{(1)}, \quad 
X_k^\mathrm{old} = (\rho X_k)^\mathrm{old} / \sum_k (\rho X_k)^\mathrm{old}, \quad 
X_k^\mathrm{new} = (\rho X_k)^\mathrm{new} / \sum_k (\rho X_k)^\mathrm{new},
\end{equation}
and finally define $A_{X_k}$ using
\begin{equation}
\frac{X^{(1)} - X^\mathrm{old}}{\Delta t}= A_{X_k}.
\end{equation}
\item {\tt species\_pred\_type} = 2 (predict $\rho'$ and $(\rho X_k)$)
\begin{equation}
I_{\rho X_k} = \frac{(\rho X_k)^\mathrm{new} - (\rho X_k)^\mathrm{old}}{\Delta t} - A_{\rho X_k}.
\label{eq:sdc:Irhoo}
\end{equation}
\item {\tt species\_pred\_type} = 3 (predict $\rho$ and $X_k$)

See {\tt species\_pred\_type} = 1 for the form of $I_{X_k}$.
\end{itemize}

@@

\begin{itemize}
\item {\tt enthalpy\_pred\_type} = {\tt predict\_rhohprime} (not implemented yet)

Here we need an $I_{\rho h}$ term for the $(\rho h)'$ evolution
equation (see Eq.~\ref{rhohprime equation}).  In this case we will use
$I_{(\rho h)'} = I_{\rho h}$

In the future, when we enable base state evolution, the base state enthalpy
evolution equation may need to know about the $I_{\rho h}$ source term.
In particular, should $(\rho h)_0$ see a $\overline{(\rho \Hnuc)}$ term?
what about an average thermal diffusion?

@@

This is the most straightforward prediction type.  The SDC solver
integrates the equation for $(\rho h)$:
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h \Ub) + \frac{Dp_0}{Dt}  + \rho H_{\rm nuc} 
\end{equation}
(shown here without diffusion or external heat sources).  Expanding
the time derivative and divergence, and using the continuity equation
we see:
\begin{equation}
\frac{\partial h}{\partial t} = -\Ub \cdot \nabla h + \frac{1}{\rho} \frac{Dp_0}{Dt}  + \frac{1}{\rho} (\rho H_{\rm nuc}) \label{eq:sdc:h}
\end{equation}
Comparing these equations, we see that 
\begin{equation}
I_{h}  = \frac{1}{\rho^{n+\myhalf}} \left [
    \frac{(\rho h)^\mathrm{new} - (\rho h)^\mathrm{old}}{\Delta t} - A_{\rho h} \right ]
\end{equation}
(Andy's Idea) Form $I_h$ in the same way we would form $I_{X_k}$ from above:
\begin{equation}
I_h = \frac{h^\mathrm{new} - h^\mathrm{old}}{\Delta t} - A_h,
\end{equation}
where we first define
\begin{equation}
\frac{(\rho h)^{(1)} - (\rho h)^\mathrm{old}}{\Delta t} = A_{\rho h},
\end{equation}
and then define $h$,
\begin{equation}
h^{(1)} = (\rho h)^{(1)} / \sum_k(\rho X_k)^{(1)}, \quad h^\mathrm{old} = (\rho h)^\mathrm{old} / \sum_k(\rho X_k)^\mathrm{old}, \quad h^\mathrm{new} = (\rho h)^\mathrm{new} / \sum_k(\rho X_k)^\mathrm{new},
\end{equation}
and finally define $A_h$ using
\begin{equation}
I_h = \frac{h^{(1)} - h^\mathrm{old}}{\Delta t} = A_h.
\end{equation}
